Promise and peril: Agnostic identification methods\\ for detecting differential item functioning

### 2020-03-19 paper/CopyOfpaper.tex
getting there

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Promise and peril: Agnostic identification methods for detecting differential item functioning},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\usepackage{float}
\usepackage[autostyle, english = american]{csquotes}
\setlength{\parindent}{4em}
\setlength{\parskip}{2em}
\usepackage{setspace}\doublespacing
\AtBeginEnvironment{tabular}{\singlespacing}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

{
\setcounter{tocdepth}{5}
\tableofcontents
}
\clearpage

\hypertarget{intro}{%
\section{Introduction}\label{intro}}

Inspired by Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}), we think of differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when an item response model with too few ability dimensions is imposed. From this perspective, the term \enquote{differential item functioning} is, perhaps, a misnomer as DIF is better thought of as a property of the student, as opposed to the item. For example, Ackerman (\protect\hyperlink{ref-ackerman1992didactic}{1992}) describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability. In this case, math ability is the \enquote{target ability,} and verbal ability is the \enquote{nuisance ability.} Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies DIF.

Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item. For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group. They fix the item easinesses for the reference group, \({b_j}^{\text{ref}}\), to values from a previous study. They set item easinesses for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits. They simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group. They generate item responses according to the Rasch model, which specifies that the probability of student \(i\) responding correctly to item \(j\) is \[P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)\] where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.

For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional. For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously \({b_j}^{\text{ref}}\) for all students. Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension. The target ability is what was previously just ability where \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group. The nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) for all students in the reference group and \({\eta_i}^{\text{ref}} = -1\) for all students in the focal group. We now need to use a compensatory 2PL model where the slope on target ability \(a_{1j} = 1\) for all items (consistent with the Rasch model) and the slope on nuisance ability \(a_{2j} = 0.6\) for all items with DIF, and \(a_{2j} = 0\) otherwise. Again, 0.6 is the magnitude of DIF in logits. According to the two-dimensional compensatory 2PL model, the probability student \(i\) responds correctly to item \(j\) is
\[
\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
\]
This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.

If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower dimensional model? Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}) tested this idea with the goal of a \enquote{more satisfying description of the secondary abilities} {[}p.~144{]}. He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that \enquote{a priori knowledge of the true factor structure} is necessary {[}p.~144{]}. It's hard to imagine how one would have such knowledge. Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters.

\hypertarget{agnostic-identification-methods}{%
\section{Agnostic identification methods}\label{agnostic-identification-methods}}

Psychometricians have long been in search of the perfect DIF detection method. It has been well demonstrated that the IRT-based likelihood ratio test (LRT) is effective at detecting DIF (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}). As a result, we avoid methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which muddies the waters by moving away from the IRT framework, and has been shown to perform no better --- and in some cases worse --- than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990}).

The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions --- about either relative group ability or which items might have DIF --- can be made. That is, we're interested in agnostic identification methods (hereafter referred to as \enquote{AI methods}) so that LRTs can be used to test items for DIF. For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection. And, in search of a coherent framework, we sometimes edit names of existing methods. We recognize that others have done the same (e.g.~Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.

In this section, we summarize existing AI methods and propose extensions of those methods. We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test. Target ability is simulated \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group. Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) and \({\eta_i}^{\text{foc}} = -1\). The slope on target ability is set to \(a_{1j} = 1\) for all items. The slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise. We can also, of course, describe these conditions from the DIF-as-item-property view where there is no nuisance ability. Instead, \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for the first five items and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5\) for the last three items.

In general, we denote the mean ability of the reference and focal group as \(\mu^\text{ref}\) and \(\mu^\text{foc}\), respectively. We follow the common and inconsequential practice of identifying the scale by setting \(\mu^\text{ref} = 0\). The fundamental challenge is to use an AI method to identify the model, which allows for disentagling the estimation of \(\hat\mu^\text{foc}\) from the estimation of \(\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\). The most common AI approach is to select a group of anchor items that are assumed to be DIF-free. These anchor items identify the model, thereby allowing for the estimation of \(\hat\mu^\text{foc}\), and the remaining items can be tested for DIF using an LRT.

Computing is done in R (R Core Team \protect\hyperlink{ref-rcore}{2019}), model fitting in the mirt R package (Chalmers \protect\hyperlink{ref-chalmers2012mirt}{2012}), and data wrangling/visualization in the suite of R packages known as the tidyverse (Wickham \protect\hyperlink{ref-tidy}{2017}). Code is available at \emph{(TODO)}.

\hypertarget{anchoritems}{%
\subsection{Anchor items}\label{anchoritems}}

\hypertarget{all-others-as-anchors-aoaa}{%
\subsubsection{All-others-as-anchors (AOAA)}\label{all-others-as-anchors-aoaa}}

Meade and Wright (\protect\hyperlink{ref-meade2012solving}{2012}) compared the most commonly used AI methods and unequivocally recommended the all-others-as-anchors (AOAA) method. AOAA tests each item for DIF one at a time using all of the other items as anchors. For example, when testing the first item for DIF, all of the other items are used as anchors. This is done using an LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups (Thissen, Steinberg, and Wainer \protect\hyperlink{ref-thissen1993detection}{1993}). Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on. The items for which the flexible model outperforms the baseline model (typically based on a \(\chi^2\) test) are identified as having DIF, and the rest of the items become anchor items. AOAA is implemented in the mirt R package, and is called by passing scheme = \enquote{drop} to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).

Edelen et al. (\protect\hyperlink{ref-edelen2006identification}{2006}) used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF. How can they be sure that its those 10 items and not the other 11 items with DIF? They cannot be. Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible. More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group. Undesirably, most applications of AI methods and many simulation studies do not make explicit the assumptions of the AI method (Strobl et al. \protect\hyperlink{ref-strobl2018anchor}{2018}). In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.

Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies (Stark, Chernyshenko, and Drasgow \protect\hyperlink{ref-stark2006detecting}{2006}). A simple thought experiment illustrates how AOAA fails: Imagine a test with a sufficiently large number of students and three items where the first item has DIF, and the other two do not. Using AOAA, all items test positive for DIF. The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated. This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until Andrich and Hagquist (\protect\hyperlink{ref-andrich2012real}{2012}) coined the term \enquote{artificial DIF.}

\hypertarget{all-others-as-anchors-all-significant-aoaa-as}{%
\paragraph{All-others-as-anchors-all-significant (AOAA-AS)}\label{all-others-as-anchors-all-significant-aoaa-as}}

One way to attempt to counter artificial DIF is with purification through iterative anchor selection. For example, Drasgow (\protect\hyperlink{ref-drasgow1987study}{1987}) started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively --- with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model --- until no more items tested positively. Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) named this technique Iterative-backward-AOAA with \enquote{backward} (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF-free. We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS). Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set. AOAA-AS is implemented in the mirt R package, and is called by passing scheme = \enquote{drop\_sequential} to the DIF function.

AOAA-AS might seem like an improvement, but it doesn't solve a fundamental problem of AOAA: What does one do when all items test positive for DIF? With a sufficient sample size and at least one item with DIF, this will necessarily be the case. In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF, and there are no anchor items. Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) encountered precisely this problem in their simulation study and chose to select a single anchor item randomly. Woods (\protect\hyperlink{ref-woods2009empirical}{2009}) suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF. It's unclear if one should proceed if those four items test positive for DIF, too.

\hypertarget{all-other-as-anchors-one-at-a-time-aoaa-oat}{%
\paragraph{All-other-as-anchors-one-at-a-time (AOAA-OAT)}\label{all-other-as-anchors-one-at-a-time-aoaa-oat}}

We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been explicitly proposed. AOAA-OAT is inspired by Hagquist and Andrich (\protect\hyperlink{ref-hagquist2017recent}{2017}), who, in general, assert that \enquote{items showing DIF initially should not be resolved simultaneously but sequentially} {[}p.~6{]}. Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the \(\chi^2\) test statistic, is removed from the anchor set. The process continues iteratively until no new items display DIF. AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF. As a result, we believe that AOAA-OAT is less likely to be \enquote{fooled} by artifical DIF. Note that AOAA-OAT is not currently implemented in the R package mirt.

Applying AOAA-OAT to our thought experiment demonstrates its effectiveness. The initial AOAA removes the real DIF item from the anchor set because it exhibits the most DIF. In the next step, both of the other items test negative for DIF, and we arrive at the correct conclusion. To work, AOAA-OAT has two requirements: First, that at least two items do not have DIF, and second, that the set of items without DIF are more homogeneous than other sets of items.

\hypertarget{summary-and-performance}{%
\paragraph{Summary and performance}\label{summary-and-performance}}

Table \ref{table:themethods} summarizes the three all-others-as-anchors methods. It's useful to remember that AOAA is not an iterative procedure. The methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the new, more conservative method. In our one-run simulation, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified. On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.

\begin{table}[H]
\caption{Summary of the three all-others-as-anchors agnostic identification methods}
\centering
\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
\toprule

@@

\bottomrule
\end{tabular}
\label{table:themethods}
\end{table}

\hypertarget{equal-means-clustering-emc}{%
\subsubsection{Equal means clustering (EMC)}\label{equal-means-clustering-emc}}

Bechger and Maris (\protect\hyperlink{ref-bechger2015statistical}{2015}) proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the \enquote{anchor cluster.} They pointed out that one way around the unidentifiability issue is to consider only relative item parameters. For each group, the relative easinesses for each pair of items can be stored in the matrix \(\mathbf{R}^{\text{ref}}\) with entries \({R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}\). The ultimate matrix of interest is \(\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}\) which is the \enquote{differences between groups in the pairwise differences in (easiness) between items} {[}p.~323{]}.

The general idea of identifying clusters of items is intriguing. However, their approach is needlessly complicated, and they did not describe a process for moving from \(\Delta \mathbf{R}\) to an anchor cluster. Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) extended their work by proposing one such process. \(\Delta \mathbf{R}\) is skew-symmetric and of rank 1, which means that all information is contained in a single row or column. Accordingly, they use k-means clustering on just the first column of \(\Delta \mathbf{R}\) where the number of clusters, k, is chosen by minimizing BIC. They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster. Unfortunately, in their simulation study, they find that BIC identifies only a single cluster, so they end up using all of the items as anchors.

We propose a new cluster-based approach, which we call \enquote{equal means clustering} (EMC). Instead of working with an arbitrary column from \(\Delta \mathbf{R}\), we work with the vector \(\tilde{\mathbf{d}}\) of differences in item easiness, \(\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}\), where the model is identified by setting \(\mu^\text{foc} = 0\) (recall \(\mu^\text{ref}\) is always set to \(0\), thus the name \enquote{equal means clustering}). As a result, all differences in performance --- either from group ability differences or DIF --- manifesting in the item easiness difference parameter \(\tilde{d_j}\). The tilde is used to denote parameters estimated with \(\mu^\text{foc}\) arbitrarily set to \(0\).

Instead of choosing k with BIC, we use the gap statistic method recommended by Hastie, Tibshirani, and Walther (\protect\hyperlink{ref-hastie2001estimating}{2001}). The largest cluster is chosen as the anchor cluster. If there is a tie for the largest cluster, the cluster with the lowest standard deviation of \(\tilde{\mathbf{d}}\) is selected. This process assumes that the largest cluster (but not necessarily the majority) of items do not contain DIF. In our one-run simulation, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.

\hypertarget{the-equal-means-multiple-imputation-logit-graph-em-milg}{%
\subsubsection{The equal means, multiple imputation logit graph (EM-MILG)}\label{the-equal-means-multiple-imputation-logit-graph-em-milg}}

AI methods are generally designed to automatically detect DIF without any human judgement. On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.

We propose a new method, the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst. Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting \(\mu^\text{foc}\) to \(0\). Again, the result is that all differences in performance manifest in \(\tilde{\mathbf{d}}\). To measure the variation in each \(\tilde{{d_j}}\), the item parameter covariance matrix is estimated using Oakes' identity (Chalmers \protect\hyperlink{ref-chalmers2018numerical}{2018}). Then, multiple imputations (MI) (Yang, Hansen, and Cai \protect\hyperlink{ref-yang2012characterizing}{2012}) are drawn to estimate the distribution of \(\tilde{d_j}\) for each item. These are the distributions displayed in a EM-MILG. The method is inspired in part by Pohl, Stets, and Carstensen (\protect\hyperlink{ref-pohl2017cluster}{2017}) who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and Talbot III (\protect\hyperlink{ref-talbot2013taking}{2013}) who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.

Figure \ref{fig:emmilg} shows the EM-MILG for our one-run simulation. We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance. The challenge, then, is to select the anchor items. The analyst might assume that --- because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits --- items 1-5 are unbiased and can be used as anchor items.

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/emmilg-1}

}

\caption{A equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.}\label{fig:emmilg}
\end{figure}

After choosing the anchor items, the model is refit. The new model is identified by setting \(d_j = 0\) for the anchor items, instead of by setting \(\mu^\text{foc} = 0\). The same process of using multiple imputations to estimate the distribution of \(\tilde{d_j}\) can be used with the new model. Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG). As expected, in our one-run simulation, selecting the first five items as anchors correctly results in \(\hat d_j \approx 0.5\) for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/milg-1}

}

\caption{A multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group. Anchor items are fixed by setting $d_j = 0$.}\label{fig:milg}
\end{figure}

One of our key concerns with typical AI methods is that they can lull the analyst into a false sense of security. Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items. A EM-MILG combats this concern by presenting all information clearly to the analyst. In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased. Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data. And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG. In particular, we could visualize DIF in the non-anchor items after using, for example, AOAA-OAT or EMC to identify the model.

\hypertarget{anchor-points}{%
\subsection{Anchor points}\label{anchor-points}}

The previously discussed AI methods select a set of anchor items, whether it is an algorithm or the analyst that makes that selection. The anchor items are used to estimate \(\hat \mu^\text{foc}\). An alternative strategy is to directly set the anchor point, \(\mu^{\star\text{foc}}\). Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, allowing all items to be tested for DIF. The question then becomes the following: How is the anchor point selected?

\hypertarget{maximizing-the-gini-index-maxgi}{%
\subsubsection{Maximizing the Gini index (MAXGI)}\label{maximizing-the-gini-index-maxgi}}

Strobl et al. (\protect\hyperlink{ref-strobl2018anchor}{2018}) suggest using the Gini index (Gini \protect\hyperlink{ref-gini1912variabilita}{1912}) to select the anchor point. The Gini index is typically used to measure the inequality of wealth distribution in a country. For example, South Africa typically has the highest Gini index of all measured countries, meaning that it is the country with the most unequal wealth distribution (Chitiga, Sekyere, and Tsoanamatsie \protect\hyperlink{ref-chitiga2015income}{2015}). In general, the Gini index \enquote{takes high values if, for example, a small minority of persons has a lot of wealth while the vast majority has very little} (Strobl et al. \protect\hyperlink{ref-strobl2018anchor}{2018}, 7).

\(\mu^{\star\text{foc}}\) is selected by maximizing the Gini index (thus the abbreviation MAXGI). The intuition and assumption is that the items without DIF are the most homogeneous. Denoting a function that calculates the Gini index from a vector of non-negative elements as \(G(\mathbf{x})\), MGI sets
\[
\mu^{\star\text{foc}} = \mathop\mathrm{arg\,max}_{\mu^\text{foc}} G(|\mu^\text{foc} + \tilde{\mathbf{d}}|)
\]
where \(\mu^\text{foc} \in (-\infty, \infty)\) and \(\mu^\text{foc}\) is added to each element of \(\mathbf{d}\).

For our one-run simulation, Figure \ref{fig:ginipath} shows the gini coefficient at a variety of possible \(\mu^\text{foc}\) values. The result of MAXGI is \(\mu^{\star\text{foc}} = -0.99\), which is extraordinarily close to the data-generating value of -1.

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/ginipath-1}

}

\caption{Maximizing the Gini index (MAXGI) to select the anchor point.}\label{fig:ginipath}
\end{figure}

The model is then refit with the identifying assumption that \(\mu^{\text{foc}} = \mu^{\star\text{foc}}\), and the results can be displayed in a MILG as is shown in Figure \ref{fig:ginimilg}.

\begin{figure}[H]

{\centering \includegraphics[width=0.7\linewidth]{paper_files/figure-latex/ginimilg-1}

}

\caption{The MILG with $\mu^{\text{foc}}$ set to $\mu^{\star\text{foc}} = -0.99$.}\label{fig:ginimilg}
\end{figure}

\hypertarget{minimizing-between-curves-minbc}{%
\subsubsection{Minimizing between curves (MINBC)}\label{minimizing-between-curves-minbc}}

Raju's area method (Raju \protect\hyperlink{ref-raju1988area}{1988}) measures the amount of DIF by calculating the area between the item characteristic curves, the function that maps the student's ability to their probability of correct response, of the two groups:
\[
\text{Area Between Curves} = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})|
\]
Raju's area method has been cited as one of the most commonly used IRT-based DIF detection methods (Magis et al. \protect\hyperlink{ref-magis2011generalized}{2011}). However, Raju's area method is not an AI method because the item characteristic curves still need to be linked by anchor items or an anchor point. An additional weakness is that the area is unweighted, so all values of \(\theta\) matter equally, despite some being much more realistic than others.

To adapt Raju's area method into an AI method, we propose a new method, which we call \enquote{minimizing the area between curves} (MINBC). To understand MINBC, imagine a scenario in which the data-generating process is \(\mu^{\text{foc}} = \mu^{\text{ref}}\) and \(d_j = 0 \forall j\), so that there is no DIF. The fundamental identification problem is that there are an infinite number of models with the same likelihood from which to choose. For example, we could correctly assume that the focal group has the same ability as the reference group and fix \(\mu^{\star\text{foc}} = 0\). The model would then estimate \(\hat d_j \approx 0 \forall j\), and we would correctly conclude the groups have the same ability and there is no DIF. Alternatively, we could assume that the focal group has \(\mu^{\star\text{foc}} = 3\). The model would then compensate by finding \(\hat d_j \approx -3 \forall j\), and we would incorrectly conclude that the focal group is high ability, but every item contains DIF against them. Both of these models have the same likelihood, so how should one choose which model to believe? MINBC chooses the model with the least amount of total DIF, as measured by the total weighted area between the item characteristic curves. As a result, the likelihood tie is broken by preferring to explain differences in performance across groups by ability differences (as opposed to DIF).

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Promise and peril: Agnostic identification methods for detecting differential item functioning},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{lineno}
\linenumbers
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\usepackage{float}
\usepackage[autostyle, english = american]{csquotes}
\setlength{\parindent}{4em}
\setlength{\parskip}{2em}
\usepackage{setspace}\doublespacing
\AtBeginEnvironment{tabular}{\singlespacing}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

{
\setcounter{tocdepth}{5}
\tableofcontents
}
\begin{table}[H]
\caption{Summary of agnostic identification methods}
\centering
\begin{tabular}{|p{4cm}|p{6cm}|p{4cm}|}
\toprule

@@

### 2020-03-19 paper/CopyOfpaper_2a_anchor_items.Rmd
getting there

@@

### The equal means, multiple imputation logit graph (EM-MILG)

AI methods are generally designed to automatically detect DIF without any human judgement. On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.

>>>>

## The equal means, multiple imputation logit graph (EM-MILG)

AI methods are generally designed to automatically detect DIF without any human judgement. On the other hand, it can be useful to present information to the analyst in a way that empowers them to visualize their data.

@@
